Covid-19 A Spatio-Temporal Investigation.

### 2021-01-08 Notebooks/paper_covid.rmd
Update paper_covid.rmd

 O objetivo deste trabalho será portanto traçar um perfil sobre o comportamento espacial da pandemia causada pela Covid-19 ao utilizar de técnicas estatísticas para quantificação da dependência espacial como o cálculo do índice de Moran global e local $\!\!^4$, além de uma estatística de scan para identificação de *clusters* emergentes do número de casos no espaço-tempo, bem como analisar as políticas adotadas por cada país e o comportamento da pandemia ao longo do tempo.

>>>>

 The main goal of this paper is to present methods used to identify spatial and spatio-temporal dependence in the number of infections and deaths caused by COVID-19, so they can be used in future or hypothetical situations where is necessary to investigate the behavior of an pathogen agent in space and time. The spatial method was calculated by using the Global Moran's I and the Local Moran's I $\!\!^4$, which is also known as a LISA statistic (Local Indicator of Spatial Association), both being used as indicators of spatial association. Spatio-temporal association was evaluated by computing a spatio-temporal poisson scan statistic, proposed by Kulldorff (place reference here) to identify emerging or "alive" *clusters* in space and time.
